\subsection*{Learning Objectives}
	Students need to be able to\ldots
		\begin{itemize}\it 
			\item Model with ODEs. \\[-20pt]
			\item Analyze equilibria and their stability. \\[-20pt]
			\item Model with PDEs.
		\end{itemize}


%	
\subsection*{Context}
	
In lecture we studied some examples of finding the stability of equilibrium solutions as well as some modelling with PDEs, both the transport equation and the traffic model.

In this tutorial, we will continue practicing studying the stability of the equilibrium solution as well as studying a model given by an ODE and then we will delve deeper into the traffic model.

%\subsection*{Resources for TAs}



\subsection*{Before Tutorial}


The midterm is next week on Friday (during lecture), so:
\begin{itemize}
	\item Make a ``when to meet'' to decide on the best time for 2h office hours before the test on Wednesday (after 2pm) or Thursday anytime.
	\item Let Bernardo know the date+time chosen so it can be included in the midterm info announcement on Friday.
\end{itemize}


Send an announcement to students letting them know that they will \textbf{NOT} need to bring a laptop. % and will be using Jupyter Notebooks \url{https://utoronto.syzygy.ca}.

Students should review how to find the stability of an equilibrium solution of an ODE as well as how to sketch a phase portrait of an ODE (both a single ODE and a system of ODEs).


\subsection*{What to Do}
	
Introduce the learning objectives for the day's tutorial. \\

%Have students get into small groups and start on \#1 -- each group needs to have at least 1 laptop. \\





%\subsection*{Notes}
%
%	\begin{enumerate}
%		\item Note
%	\end{enumerate}
	
	

	
	
