		\begin{objectives}
			In this tutorial you will explore some modelling towards optimization problems and using the techniques we have learned to analyze the results.

				These problems relate to the following course learning objectives:
				\begin{itemize}\it 
					\item Model optimization problems. \\[-20pt]
					\item Recognize what type of optimization problem they need. \\[-20pt]
					\item Use Jupyter Notebook to approximate the solution. \\[-20pt]
					\item Create a visualization of the problem and the solution.
				\end{itemize}
%						\textit{Apply linear algebra techniques to classify solutions of linear systems of ordinary differential
%			equations including rigorously classifying the stability of equilibrium solutions and creating
%			linear approximations to non-linear systems of ordinary differential equations.}
		\end{objectives}

\vspace{-.5em}
\subsection*{Problems}
\vspace{-.5em}


\begin{enumerate}
	\item\label{q1} You recently started playing board games and your board game collection is getting large enough that you decided to buy a bookshelf to store your games. 
	
	You don't have a lot of money, so you find a cheap used $2 \times 2$ Ikea Kallax bookshelf.
	The dimensions of each shelf are 35cm$\times$35cm$\times$35cm.
	
	Your board game collection consists of 19 games and their boxes have a square profile 30cm$\times$30cm with the following thicknesses:
	\begin{itemize}
		\item 5 games with box thickness of 7cm
		\item 10 games with box thickness of 8cm
		\item 4 games with box thickness of 9cm
	\end{itemize}

	\begin{enumerate}
		\item What is the optimal way to place them in your new bookshelf?
		\item Create a visualization of the problem and solution.

		\item Does the solution change if one of the shelves came with a defect and was 36cm wide instead?
	\end{enumerate}


\vfil 

\item \label{q2} You live in a condo building and want to \textbf{optimize} the resting position of the elevators.

	Here are some details about the building:
	\begin{itemize}
		\item There are 3 elevators
		\item The building has 30 floors with the ground floor on floor 0 and apartments on floors 1--30 (inclusive).
		\item There are 10 apartments per floor with 2 people living in each apartment.
		\item The elevators take 1 second to open/close the doors
	\end{itemize}

	\begin{enumerate}
		\item Use the DABAR framework to solve this problem.
		\item Create a visualization of the problem and solution.
		\item Calculate the sensitivity with respect to the number of people in each floor and with respect to the time it takes the elevators to open/close the doors
	\end{enumerate}
	
	

	
	



	
	
	
	
	
\end{enumerate}

















